Wavefield Reconstruction Inversion: an example

### 2020-02-23 repro/papers/wri1d/paper.tex
near final

The main result of this paper is that in one simple case, in which all
of the necessary computations can be carried out by hand, this hope is
not realized: minimization of the WRI objective is just as likely to
be trapped in a spurious local minimizer as is the FWI objective.  The
context of this conclusion is a simple transmission inverse problem
for the 1D acoustic wave system, which models pulse transmission along
a 1D continuum from a source point to a receiver point. The pulses
used in this thought experiment are short, so the main information
content of the data is the time of transit from source to
receiver. The predominant information about the material model, in
this case reduced to the wave velocity (a scalar function of
position), is just this transit time, so I constrain both the target
wave velocity $c_*$ generating the data and the trial wave velocity
$c$ to be constant, that is, independent of position along the 1D
continuum. I introduce a family of inverse problems, parametrized by a
parameter $\lambda$ playing the role of wavelength. For sufficiently small
$\lambda$, it is possible to show explicitly that the FWI problem
possesses local minimizers far from the global minimizer at the target
$c_*$, and that initiating a local iterative optimization, such as
steepest descent or Newton's method, at a distance from $c_*$ bounded
below by a multiple of the ``wavelength'' $\lambda$ will result in
convergence to these spurious local minima. That is, FWI 
behaves in exactly the manner described in much of the literature on
this topic. However, an analysis of WRI applied to the same context
yields the same result: spurious local minima exist for sufficiently
small $\lambda$, and will be found by local optimization unless the
starting point is within $O(\lambda)$ of the target. That is, WRI
behaves in a manner qualitatively indistinguishable from FWI. In
particular, its ability to allow good fit to data for small $\alpha$
does not safeguard it from failure to converge globally to a ``good''
local minimum.

This paper begins with a description of the 1D inverse transmission
problem and explicit computation of various components of the FWI
approach, based on explicit solution of the 1D acoustic system
presented in Appendix A. In the third section I introduce the $\lambda-$dependent
family of problems, and establish the asymptotic properties of FWI as
$\lambda \rightarrow 0$. The fourth section develops the algebraic
structure of WRI, culminating in a remarkable identity revealing WRI
to be equivalent to minimization of a weighted norm of the data
residual, with a weight operator depending on the coefficient vector
$c$. This result is quite general, applying to essentially any
realization of WRI, and already shows that its behaviour must be
closely related to that of FWI. In the fifth section I return to 
1D acoustics problem and the $\lambda$-dependent family of inverse
problems, compute that weight operator explicitly, and deduce the
global behaviour of WRI in this instance. The paper ends with a
discussion of the relation of the analysis presented here to a
previous analysis of optimization formulations of wave inversion
problems based on parameter-dependent quadratic forms
\cite[]{StolkSymes:03}, and a brief discussion of some other
alternatives to FWI.

@@

The formulation of the inverse problem via least-squares requires a choice of Hilbert space structure for the domain and range of $S[c]$. A natural choice is
\begin{equation}
  \label{eqn:defmoddom}
  S[c]: L^2([z_{\rm min},z_{\rm max}] \times \bR) \rightarrow L^2[0,T]
\end{equation}
That is, the support of f will be assumed to lie in the strip $[z_{\rm min},z_{\rm max}] \times \bR$.
As it will turn out, the choice of spatial interval $[z_{\rm min},z_{\rm max}]$ is arbitrary, so long as it has positive length and contains $z_s$ in its interior.

@@

The version of FWI discussed here presumes that the source field
corresponding to the data $d$ is supported at a point $z_s \ne z_r$,
and is known. This field depends on a function of time (``wavelet'')
$w \in L^2(\bR)$. Formally, the resulting acoustic field satisfies the
system \ref{eqn:awe1d} with $f=w(t)\delta(z-z_s)$. Inserting this
expression in the explicit expression \ref{eqn:defmodhom}, obtain
\begin{equation}
  \label{eqn:defmodpt}
  S[c](w\delta(\cdot-z_s)) = \frac{1}{2c}w\left(t - \frac{|z_r-z_s|}{c}\right) = S_p[c]w(t)
\end{equation}
Thus for this restricted class of source fields, the FWI objective can be redefined as
\begin{equation}
  \label{eqn:deffwipt}
  J_{\rm FWI}[c,d.w] = \frac{1}{2}\|S_p[c]w -d\|^2.
\end{equation}
$S_p[c]$ is a bounded operator with domain $L^2(\bR)$ and range $L^2([0,T])$.

To end this section, it is necessary to address an irritating technical point: the point source defined above is not a member of the domain of $S[c]$, as it was defined in display \ref{eqn:defmoddom}, so the left-hand side of equation \ref{eqn:defmodpt} does not actually make sense. The fix for this incompatibility actually elucidates the relation between $S[c]$ and $S_p[c]$. Appendix B describes the construction of a family of bounded injective operators  $:L^2(\bR) \rightarrow L^2([z_{\rm min},z_{\rm max}] \times \bR)$ for which
\begin{equation}
  \label{eqn:ext}
  S_p[c] = S[c] \circ E[c]
\end{equation}
This relation exhibits $S[c]$ as an {\em extension} of $S_p[c]$ as
described by \cite{Symes:09}, and identifies WRI as an extended
inversion method.

@@

\section{Global Asymptotics of FWI}
The FWI objective is well-known to exhibit non-convexity unless
data frequency content is limited to a small range near 0 Hz, how
small being determined by other scales and by the extent to which the
initial wave velocity differs from the target.

To understand the non-convexity phenomenon and the
relation of the various scales, it is advantageous to introduce a
family of source wavelets, depending on a parameter $\lambda$, having
dimensions of time and playing
the role of wavelength:
\begin{equation}
  \label{eqn:wfam}
  \wl(t) = \frac{1}{\sqrt{\lambda}}w_1\left(\frac{t}{\lambda}\right).
\end{equation}
In the definition \ref{eqn:wfam}, the ``mother wavelet'' $w_1\in
C_0^{\infty}(0,1)$, and the scaling is chosen so fhat
$\|\wl\|_{L^2(\bR)}$ is independent of $\lambda>0$.

@@

Choose a target wave velocity $c_* \in [c_{\rm min},c_{\rm
  max}]$. Introduce a family of consisten data $\dl$, generated by
$c_*$ and the wavelet family $\wl$:
\begin{equation}
  \label{eqn:dfam}
  \dl = S_p[c_*]\wl
\end{equation}
and a corresponding family of FWI objectives:
\[
  J_{\rm FWI}[c,\dl,\wl] = \frac{1}{2}\|\dl-S_p[c]\wl\|^2
\]
\begin{equation}
  \label{eqn:fwiexpl}
  =
  \int_0^T\,dt\,\left|\frac{1}{2c_*}\wl\left(t-\frac{z_s-z_r}{c_*}\right)
    - \frac{1}{2c}\wl\left(t-\frac{z_s-z_r}{c}\right)\right|^2
\end{equation}
Note that $\mbox{supp }\wl \subset [0,\lambda]$, so
\begin{equation}
  \label{eqn:wlsupps}
  \mbox{supp }S_p[c]\wl \subset \left[\frac{z_s-z_r}{c},\lambda
    +\frac{z_s-z_r}{c}\right] \cap [0,T]
\end{equation}
The transit time condition \ref{eqn:time} implies that there exists
$\lambda_0>0$ so that for $\lambda < \lambda_0$,
\[
  \lambda +\frac{z_s-z_r}{c} < T
\]
for all admissible $c$. That is, 
for $\lambda <\lambda_0$, $\mbox{supp }S_p[c]\wl \subset(0,T)\mbox{
  for }c \in [c_{\rm min},c_{\rm max}]$, and 
\begin{equation}
  \label{eqn:constnorm}
  \|S_p[c]\wl\|^2 = \frac{1}{4c^2}\int_{-\infty}^{\infty}\,dt\,\left|
    \wl\left(t-\frac{z_s-z_r}{c}\right)\right|^2 =
  \frac{\|w_1\|^2}{4c^2}.
\end{equation}

@@

\cite{LeeuwenHerrmann:16} used the variable projection method
\cite[]{GolubPereyra:03}, eliminating the source $f$ in the inner step
and updating the bulk modulus (or an equivalent quantity) in the outer
step. That is, the problem \ref{eqn:defwri} is equivalent to
minimization of
\[
  J_{\rm WRI}[c,d,q] =
  \mbox{min}_f \frac{1}{2}(\| d-S[c]f\|^2+\alpha^2\|f-q\|^2)
\]
over $c$. This is the approach that I shall pursue here.

It will turn out to be convenient to define the residual with the
target source $q$ as $r[c]=d-S[c]q$, and set $g=f-q$. Then this
definition can be rewritten as
\begin{equation}
\label{eqn:defvpm}
  J_{\rm WRI}[c,d,q] =\mbox{min}_g \frac{1}{2}(\| r[c]-S[c]g\|^2+\alpha^2\|g\|^2)
\end{equation}

@@

``Most'' source fields $f$ are non-radiating, that is, $S[c]f=0$, and
such sources contribute nothing to the data fit term in the definition
of $J_{\rm WRI}$. If the domain and range of $S$ were finite
dimensional (which of course they are, after discretization), then the
Fundamental Theorem of Linear Algebra identifies the null space of
$S[c]$ (the non-radiating sources) as the orthocomplement of the range
of the transpose $S[c]^T$ \cite[]{Strang:93}. In the
infinite-dimensional setting of this paper, the Closed Range Theorem
\cite[]{Yosida} states that the same is true if the range of $S[c]$ is
closed. There are various ways to ensure this property, but the
simplest is relevant here:
\begin{quote}
  For all admissible models $c$, $S[c]$ is surjective.
\end{quote}
That is, {em any data can be fit exactly using the extended model
  space (the domain of $S$)}.
This property is a characteristic of of extended modeling
methods \cite[]{geoprosp:2008}: the ability to fit any data appears to
be a essential for such methods to produce objectives
without spurious local minima. It holds for the problems for which WRI
has been advocated. For the simple model problem considered
here, surjectivity follows from the explicit expression for $S[c]f$,
as was noted in the discussion following equation \ref{eqn:defmodhom}.

Assuming that $S[c]$ is surjective for any admissible $c$, the
orthocomplement of the subspace of non-radiating sources is the range
of the adjoint operator $S[c]^T$. In the
definition \ref{eqn:defvpm}, decompose $g = S[c]^Te + n$, in
which $e$ is the same type of object as $d$ and $S[c]n=0$ (that is,
$n$ is a non-radiating source), and note that the decomposition is
orthogonal. Then
\[
  J_{\rm WRI}[c,d,q] =
  \mbox{min}_{e,n} \frac{1}{2}(\| d-S[c](S[c]^Te+q)\|^2+\alpha^2(\|S[c]^Te\|^2+\|n\|^2))
  \]
\begin{equation}
  \label{eqn:defvpmred}
   \mbox{min}_{e} \frac{1}{2}(\|r[c]-S[c]S[c]^Te\|^2+\alpha^2\|S[c]^Te\|^2).
 \end{equation}

This reformulation has some computational advantages \cite[]{WangYingst:SEG16,Herrmann:SEG19}, but also leads to a useful analytic transformation of the WRI problem. The minimizer on the RHS of equation \ref{eqn:defvpmred} is the solution $e=e_{\alpha}[c]$ of the normal equation
\[
 ( (S[c]S[c]^T)^2 + \alpha^2S[c](S[c]^T)e = S[c]S[c]^Tr[c]
\]
whence
\[
  S[c]S[c]^Te_{\alpha}[c] = S[c]S[c]^T(S[c]S[c]^T+\alpha^2I)^{-1}r[c]
\]
Since the null space of $S[c]$ is orthogonal to the range of $S[c]^T$
under the surjectivity assumption, $S[c]S[c]^T$ is injective, whence
\begin{equation}
  \label{eqn:norsol}
  e_{\alpha}[c]=(S[c]S[c]^T+\alpha^2I)^{-1}r[c]
\end{equation}
Consequently
\[
  J_{\rm WRI}[c,d,q] =
  \frac{1}{2}(\|r[c]-S[c]S[c]^Te_{\alpha}[c]\|^2+\alpha^2\|S[c]^Te_{\alpha}[c]\|^2)
\]
\[
  = \frac{1}{2}\left(\|r[c]-S[c]S[c]^T
    (S[c]S[c]^T+\alpha^2I)^{-1}r[c]\|^2 \right.
\]
\[
  \left.+\langle (S[c]S[c]^T+\alpha^2I)^{-1}r[c],
  S[c]S[c]^T (S[c]S[c]^T+\alpha^2I)^{-1}r[c]
  \rangle\right)
\]
\[
  =\frac{1}{2}\left(\|\alpha^2 (S[c]S[c]^T+\alpha^2I)^{-1}r[c]\|^2\right.
\]
\[
  \left.
  +\alpha^2\langle (S[c]S[c]^T+\alpha^2I)^{-1}r[c],
  -\alpha^2 (S[c]S[c]^T+\alpha^2I)^{-1}r[c]\rangle\right.
\]
\[
  + \left.\alpha^2\langle (S[c]S[c]^T+\alpha^2I)^{-1}r[c],r[c]\rangle\right)
\]
\[
  =  \frac{\alpha^2}{2}
\langle (S[c]S[c]^T+\alpha^2I)^{-1}r[c],r[c]\rangle 
\]
That is,
\begin{equation}
  \label{eqn:defwrialt}
  J_{\rm WRI}[c,d,q] = \frac{1}{2}\langle r[c], W_{\alpha}[c] r[c]\rangle
\end{equation}
with
\begin{equation}
  \label{eqn:defwriwt}
  W_{\alpha}[c] = \frac{\alpha^2}{2}(S[c]S[c]^T+\alpha^2I)^{-1}
\end{equation}
This remarkable identity shows that the WRI objective function is a {\em weighted norm of the data residual $r[c]$}.

\section{WRI for 1D Acoustics}
The preceding section provides the necessary ingredients for an assessment of the relation between WRI and FWI. While the conclusion reached below applies to many wave propagation settings, the 1D acoustic setting is particularly simple and yet illustrates clearly the nature of this relation.

@@

Next, suppose that $q = w\delta(z-z_s)$, that is, the target source is
a point source, so that $S[c]q = S_p[c]w$ in the notation used in
the discussion of FWI. Thus \ref{eqn:defwrialt} can be re-written as
\begin{equation}
  \label{eqn:wriscrewed}
  J_{\rm WRI}[c,d,q]=u[c]J_{\rm FWI}[c,d,w]
\end{equation}

@@

Define
\begin{equation}
  \label{eqn:defbeta}
\beta = \frac{z_{\rm max}-z_{\rm min}}{c_*^2} - 4 \alpha^2
\end{equation}
\begin{theorem}
  \label{thm:thm2}
  For $L$ as defined in \ref{eqn:defL}, and $\lambda <
  \lambda_0$, the minimizer of $J_{\rm WRI}[c,\dl,\wl \delta(\cdot-z_s)]$ on the
  complement of $[c_*-L\lambda, c_*+L\lambda]$ is
  \begin{itemize}
  \item $c=c_{\rm max}$ if $\beta<0$;
  \item $c=c_{\rm min}$ if $\beta>0$;
  \item any $c < c_*-L\lambda$ or $>c_*+L\lambda$ if $\beta=0$.
  \end{itemize}
\end{theorem}
\begin{proof}
  From \ref{eqn:wriscrewed}, \ref{eqn:wriwthom}, and
  \ref{eqn:fwiscrewed}, if $|c-c_*| > L\lambda$,
  \[
    J_{\rm WRI}[c,\dl,\wl\delta(\cdot-z_s)] =
    \frac{\alpha^2}{2}\left(\frac{z_{\rm max}-z_{\rm
          min}}{4c^2}+\alpha^2\right)^{-1}\frac{1}{2}\left(\frac{1}{4c^2}+\frac{1}{4c_*^2}\right)\|w_1\|^2
  \]
  \[
    =
    \frac{\alpha^2}{4}\frac{1+\frac{c^2}{c_*^2}}{z_{\rm 
        max}-z_{\rm min}+4c^2\alpha^2}
  \]
  This linear fractional function of $c^2$ is increasing, decreasing,
  or constant if $\beta>0, \beta<0$ or $\beta=0$, respectively.
\end{proof}

In other words, $J_{\rm WRI}$ has local minima far from the target
velocity $c_*$, in the same way as does $J_{\rm FWI}$. One of the
local minima will be the result of a local optimization almost surely,
unless the initial estimate of $c$ is ``within a wavelength'' of the
target velocity.

\section{Discussion}
Theorems \ref{thm:thm1} and \ref{thm:thm2} call out the chief
conclusions of this work: that at least for the 1D acoustic
transmission inverse problem, both $J_{\rm FWI}[c,d,w]$ and
$J_{\rm WRI}[c,d,q]$ exhibit local minima (in $c$) far from the global
minimum for consistent data, and the domain of attraction of the
global minimizer can be arbitrarily small. These are striking
conclusions, but the 1D acoustic transmission problem is very special
and lacks fidelity to field practice. I shall show how these
approaches to solving this special problem share properties with a
much larger family of inversion methods. The theory developed to
explain these properties suggests methods that may not suffer the
non-convexity of FWI and WRI, and in fact do not in several cases that
I will mention.

@@

One step in this construction involve buidling a constitutive defect
(pressure) source that is equivalent, in the sense of generating the
same solution, to a force (velocity) source. This construction is
presented here in the context of the 1D acoustic problem, but is a
special case of a much more general construction of considerable
interest in its own right \cite[]{BurridgeKnopoff:64}.

>>>>

The main result of this paper is that in one simple case, in which all
of the necessary computations can be carried out by hand, this hope is
not realized: minimization of the WRI objective is just as likely to
be trapped in a spurious local minimizer as is the FWI objective.  The
context of this conclusion is a simple transmission inverse problem
for the 1D acoustic wave system, which models pulse transmission along
a 1D continuum from a source point to a receiver point. The pulses
used in this thought experiment are short, so the main information
content of the data is the time of transit from source to
receiver. The predominant information about the material model, in
this case reduced to the wave velocity (a scalar function of
position), is just this transit time, so I constrain both the target
wave velocity $c_*$ generating the data and the trial wave velocity
$c$ to be constant, that is, independent of position along the 1D
continuum. I introduce a family of inverse problems, depending on a
parameter $\lambda$ playing the role of wavelength. For sufficiently small
$\lambda$, it is possible to show explicitly that the FWI problem
possesses local minimizers far from the global minimizer at the target
$c_*$, and that initiating a local iterative optimization, such as
steepest descent or Newton's method, at a distance from $c_*$ bounded
below by a multiple of the ``wavelength'' $\lambda$ will result in
convergence to these spurious local minima. That is, FWI 
behaves in exactly the manner described in much of the literature on
this topic. However, an analysis of WRI applied to the same context
yields the same result: spurious local minima exist for sufficiently
small $\lambda$, and will be found by local optimization unless the
starting point is within $O(\lambda)$ of the target. That is, WRI
behaves in a manner qualitatively indistinguishable from FWI. In
particular, its ability to allow good fit to data for small $\alpha$
does not safeguard it from failure to converge globally to a ``good''
local minimum. In fact, the $\alpha \rightarrow 0$ limit of the WRI
objective is well-defined (after scaling by $1/alpha^2$ and also behaves in
qualitatively the same way as the FWI objective. So the apparent
ability to maintain better data fit via reduction of $\alpha$ does not
lead to global behaviour asymptotically more amenable to local optimization.

This paper begins with a description of the 1D inverse transmission
problem and explicit computation of various components of the FWI
approach, based on explicit solution of the 1D acoustic system
presented in Appendix A. In the third section I introduce the $\lambda-$dependent
family of problems, and establish the asymptotic properties of FWI as
$\lambda \rightarrow 0$. The fourth section develops the algebraic
structure of WRI, culminating in a remarkable identity revealing WRI
to be equivalent to minimization of a weighted norm of the data
residual, with a weight operator depending on the coefficient vector
$c$. This identity has also been derived by \cite{Leeuwen2019note},
using a different argument. The result is quite general, applying to essentially any
realization of WRI, and already shows that its behaviour must be
closely related to that of FWI. In the fifth section I return to 
1D acoustics problem and the $\lambda$-dependent family of inverse
problems, compute that weight operator explicitly, and deduce the
global behaviour of WRI in this instance. The paper ends with a
discussion of the relation of the analysis presented here to a
previous analysis of optimization formulations of wave inversion
problems based on parameter-dependent quadratic forms
\cite[]{StolkSymes:03}, and a brief discussion of some other
alternatives to FWI.

@@

The formulation of the inverse problem via least-squares requires a choice of Hilbert space structure for the domain and range of $S[c]$. A natural choice is
\begin{equation}
  \label{eqn:defmoddom}
  S[c]: L^2([z_{\rm min},z_{\rm max}] \times \bR) \rightarrow L^2[0,T]
\end{equation}
That is, the support of f will be assumed to lie in the strip $[z_{\rm min},z_{\rm max}] \times \bR$.
As it will turn out, the choice of spatial interval $[z_{\rm min},z_{\rm max}]$ is arbitrary, so long as it has positive length.

@@

The version of FWI discussed here presumes that the source field
corresponding to the data $d$ is supported at a point $z_s \in \bR$,
and is known. The source position $z_s$ must satisfy $z_{\rm min} \le
z_s \le z_{\rm max}$ and $z_s \ne z_r$, but is otherwise arbitrary.

This point source field depends on a function of time (``wavelet'')
$w \in L^2(\bR)$. Formally, the resulting acoustic field satisfies the
system \ref{eqn:awe1d} with $f=w(t)\delta(z-z_s)$. Inserting this
expression in the explicit expression \ref{eqn:defmodhom}, obtain
\begin{equation}
  \label{eqn:defmodpt}
  S[c](w\delta(\cdot-z_s)) = \frac{1}{2c}w\left(t - \frac{|z_r-z_s|}{c}\right) = S_p[c]w(t)
\end{equation}
Thus for this restricted class of source fields, the FWI objective can be redefined as
\begin{equation}
  \label{eqn:deffwipt}
  J_{\rm FWI}[c,d.w] = \frac{1}{2}\|S_p[c]w -d\|^2.
\end{equation}
$S_p[c]$ is a bounded operator with domain $L^2(\bR)$ and range $L^2([0,T])$.

To end this section, it is necessary to address an irritating
technical point: the point source defined above is not a member of the
domain of $S[c]$, as it was defined in display \ref{eqn:defmoddom}, so
the left-hand side of equation \ref{eqn:defmodpt} does not actually
make sense. The fix for this incompatibility actually elucidates the
relation between $S[c]$ and $S_p[c]$. Appendix B describes the
construction of a family of bounded injective operators  $E[c]: L^2(\bR) \rightarrow L^2([z_{\rm min},z_{\rm max}] \times \bR)$ for which
\begin{equation}
  \label{eqn:ext}
  S_p[c] = S[c] \circ E[c]
\end{equation}
This relation exhibits $S[c]$ as an {\em extension} of $S_p[c]$ as
described by \cite{Symes:09}. WRI for 1D acoustic transmission is
based on $S[c]$, and so is identified as an extended version of FWI.

@@

\section{Global Asymptotics of 1D transmission FWI}
The FWI objective is well-known to exhibit non-convexity unless
data frequency content is limited to a small range near 0 Hz, how
small being determined by other scales and by the extent to which the
initial wave velocity differs from the target.

To understand the non-convexity phenomenon and the
relation of the various scales, it is advantageous to introduce a
family of source wavelets, depending on a parameter $\lambda$, having
dimensions of time and playing
the role of wavelength:
\begin{equation}
  \label{eqn:wfam}
  \wl(t) = \frac{1}{\sqrt{\lambda}}w_1\left(\frac{t}{\lambda}\right).
\end{equation}
In the definition \ref{eqn:wfam}, the ``mother wavelet'' $w_1\in
C_0^{\infty}(0,1)$ has dimensionless argument, and the scaling is chosen so fhat
$\|\wl\|_{L^2(\bR)}$ is independent of $\lambda>0$.

@@

Choose a target wave velocity $c_* \in [c_{\rm min},c_{\rm
  max}]$. Introduce a family of consistent data $\dl$, generated by
$c_*$ and the wavelet family $\wl$:
\begin{equation}
  \label{eqn:dfam}
  \dl = S_p[c_*]\wl
\end{equation}
and a corresponding family of FWI objectives:
\[
  J_{\rm FWI}[c,\dl,\wl] = \frac{1}{2}\|\dl-S_p[c]\wl\|^2
\]
\begin{equation}
  \label{eqn:fwiexpl}
  =
  \int_0^T\,dt\,\left|\frac{1}{2c_*}\wl\left(t-\frac{|z_s-z_r|}{c_*}\right)
    - \frac{1}{2c}\wl\left(t-\frac{|z_s-z_r|}{c}\right)\right|^2
\end{equation}
Note that $\mbox{supp }\wl \subset [0,\lambda]$, so
\begin{equation}
  \label{eqn:wlsupps}
  \mbox{supp }S_p[c]\wl \subset \left[\frac{|z_s-z_r|}{c},\lambda
    +\frac{|z_s-z_r|}{c}\right] \cap [0,T]
\end{equation}
The transit time condition \ref{eqn:time} implies that there exists
$\lambda_0>0$ so that for $\lambda < \lambda_0$,
\[
  \lambda +\frac{|z_s-z_r|}{c} < T
\]
for all admissible $c$. That is, 
for $\lambda <\lambda_0$, $\mbox{supp }S_p[c]\wl \subset(0,T)\mbox{
  for }c \in [c_{\rm min},c_{\rm max}]$, and 
\begin{equation}
  \label{eqn:constnorm}
  \|S_p[c]\wl\|^2 = \frac{1}{4c^2}\int_{-\infty}^{\infty}\,dt\,\left|
    \wl\left(t-\frac{|z_s-z_r|}{c}\right)\right|^2 =
  \frac{\|w_1\|^2}{4c^2}.
\end{equation}

@@

\cite{LeeuwenHerrmann:16} used the variable projection method
\cite[]{GolubPereyra:03}, eliminating the source $f$ in the inner step
and updating the bulk modulus (or an equivalent quantity) in the outer
step. That is, the problem \ref{eqn:defwri} is equivalent to
minimization of
\[
  J_{\rm WRI}^{\alpha}[c,d,q] =
  \mbox{min}_f \frac{1}{2}(\| d-S[c]f\|^2+\alpha^2\|f-q\|^2)
\]
over $c$. This is the approach that I shall pursue here.

It will turn out to be convenient to define the residual with the
target source $q$ as $r[c]=d-S[c]q$, and set $g=f-q$. Then this
definition can be rewritten as
\begin{equation}
\label{eqn:defvpm}
  J_{\rm WRI}^{\alpha}[c,d,q] =\mbox{min}_g \frac{1}{2}(\| r[c]-S[c]g\|^2+\alpha^2\|g\|^2)
\end{equation}

@@

``Most'' source fields $f$ are non-radiating, that is, $S[c]f=0$, and
such sources contribute nothing to the data fit term in the definition
of $J_{\rm WRI}^{\alpha}$. If the domain and range of $S$ were finite
dimensional (which of course they are, after discretization), then the
Fundamental Theorem of Linear Algebra identifies the null space of
$S[c]$ (the non-radiating sources) as the orthocomplement of the range
of the transpose $S[c]^T$ \cite[]{Strang:93}. In the
infinite-dimensional setting of this paper, the Closed Range Theorem
\cite[]{Yosida} states that the same is true if the range of $S[c]$ is
closed. There are various ways to ensure this property, but the
simplest is relevant here:
\begin{quote}
  For all admissible models $c$, $S[c]$ is surjective.
\end{quote}
That is, {em any data can be fit exactly using the extended model
  space (the domain of $S$)}.
This property is a characteristic of of extended modeling
methods \cite[]{geoprosp:2008}: the ability to fit any data appears to
be a essential for such methods to produce objectives
without spurious local minima. It holds for the problems for which WRI
has been advocated. For the simple model problem considered
here, surjectivity follows from the explicit expression for $S[c]f$,
as was noted in the discussion following equation \ref{eqn:defmodhom}.

Assuming that $S[c]$ is surjective for any admissible $c$, the
orthocomplement of the subspace of non-radiating sources is the range
of the adjoint operator $S[c]^T$. In the
definition \ref{eqn:defvpm}, decompose $g = S[c]^Te + n$, in
which $e$ is the same type of object as $d$ and $S[c]n=0$ (that is,
$n$ is a non-radiating source), and note that the decomposition is
orthogonal. Then
\[
  J_{\rm WRI}^{\alpha}[c,d,q] =
  \mbox{min}_{e,n} \frac{1}{2}(\| d-S[c](S[c]^Te+q)\|^2+\alpha^2(\|S[c]^Te\|^2+\|n\|^2))
  \]
\begin{equation}
  \label{eqn:defvpmred}
 =  \mbox{min}_{e} \frac{1}{2}(\|r[c]-S[c]S[c]^Te\|^2+\alpha^2\|S[c]^Te\|^2).
 \end{equation}

This reformulation has some computational advantages \cite[]{WangYingst:SEG16,Herrmann:SEG19}, but also leads to a useful analytic transformation of the WRI problem. The minimizer on the RHS of equation \ref{eqn:defvpmred} is the solution $e=e_{\alpha}[c]$ of the normal equation
\[
 ( (S[c]S[c]^T)^2 + \alpha^2S[c](S[c]^T)e = S[c]S[c]^Tr[c]
\]
whence
\[
  S[c]S[c]^Te_{\alpha}[c] = S[c]S[c]^T(S[c]S[c]^T+\alpha^2I)^{-1}r[c]
\]
Since the null space of $S[c]$ is orthogonal to the range of $S[c]^T$
under the surjectivity assumption, $S[c]S[c]^T$ is injective, whence
\begin{equation}
  \label{eqn:norsol}
  e_{\alpha}[c]=(S[c]S[c]^T+\alpha^2I)^{-1}r[c]
\end{equation}
Consequently
\[
  J_{\rm WRI}^{\alpha}[c,d,q] =
  \frac{1}{2}(\|r[c]-S[c]S[c]^Te_{\alpha}[c]\|^2+\alpha^2\|S[c]^Te_{\alpha}[c]\|^2)
\]
\[
  = \frac{1}{2}\left(\|r[c]-S[c]S[c]^T
    (S[c]S[c]^T+\alpha^2I)^{-1}r[c]\|^2 \right.
\]
\[
  \left.+\langle (S[c]S[c]^T+\alpha^2I)^{-1}r[c],
  S[c]S[c]^T (S[c]S[c]^T+\alpha^2I)^{-1}r[c]
  \rangle\right)
\]
\[
  =\frac{1}{2}\left(\|\alpha^2 (S[c]S[c]^T+\alpha^2I)^{-1}r[c]\|^2\right.
\]
\[
  \left.
  +\alpha^2\langle (S[c]S[c]^T+\alpha^2I)^{-1}r[c],
  -\alpha^2 (S[c]S[c]^T+\alpha^2I)^{-1}r[c]\rangle\right.
\]
\[
  + \left.\alpha^2\langle (S[c]S[c]^T+\alpha^2I)^{-1}r[c],r[c]\rangle\right)
\]
\[
  =  \frac{\alpha^2}{2}
\langle (S[c]S[c]^T+\alpha^2I)^{-1}r[c],r[c]\rangle 
\]
That is,
\begin{equation}
  \label{eqn:defwrialt}
  J_{\rm WRI}^{\alpha}[c,d,q] = \frac{1}{2}\langle r[c], W_{\alpha}[c] r[c]\rangle
\end{equation}
with
\begin{equation}
  \label{eqn:defwriwt}
  W_{\alpha}[c] = \frac{\alpha^2}{2}(S[c]S[c]^T+\alpha^2I)^{-1}
\end{equation}
This remarkable identity shows that the WRI objective function is a
{\em weighted norm of the data residual $r[c]$}.

\cite{Leeuwen2019note} gives a different derivation of an identity
equivalent to equations \ref{eqn:defwrialt}, \ref{eqn:defwriwt}.

\section{Global Asymptotics of 1D transmission WRI}
The preceding section provides the necessary ingredients for an assessment of the relation between WRI and FWI. While the conclusion reached below applies to many wave propagation settings, the 1D acoustic setting is particularly simple and yet illustrates clearly the nature of this relation.

@@

Next, suppose that $q = w\delta(z-z_s)$, that is, the target source is
a point source, so that $S[c]q = S_p[c]w$ in the notation used in
the discussion of FWI. Thus \ref{eqn:defwrialt} can be re-written as
\begin{equation}
  \label{eqn:wriscrewed}
  J_{\rm WRI}^{\alpha}[c,d,q]=u[c]J_{\rm FWI}[c,d,w]
\end{equation}

@@

Define
\begin{equation}
  \label{eqn:defbeta}
\beta = \frac{z_{\rm max}-z_{\rm min}}{c_*^2} - 4 \alpha^2
\end{equation}
\begin{theorem}
  \label{thm:thm2}
  For $L$ as defined in \ref{eqn:defL}, and $\lambda <
  \lambda_0$, the minimizer of $J_{\rm WRI}^{\alpha}[c,\dl,\wl \delta(\cdot-z_s)]$ on the
  complement of $[c_*-L\lambda, c_*+L\lambda]$ is
  \begin{itemize}
  \item $c=c_{\rm max}$ if $\beta<0$;
  \item $c=c_{\rm min}$ if $\beta>0$;
  \item any $c < c_*-L\lambda$ or $>c_*+L\lambda$ if $\beta=0$.
  \end{itemize}
\end{theorem}
\begin{proof}
  From \ref{eqn:wriscrewed}, \ref{eqn:wriwthom}, and
  \ref{eqn:fwiscrewed}, if $|c-c_*| > L\lambda$,
  \[
    J_{\rm WRI}^{\alpha}[c,\dl,\wl\delta(\cdot-z_s)] =
    \frac{\alpha^2}{2}\left(\frac{z_{\rm max}-z_{\rm
          min}}{4c^2}+\alpha^2\right)^{-1}\frac{1}{2}\left(\frac{1}{4c^2}+\frac{1}{4c_*^2}\right)\|w_1\|^2
  \]
  \[
    =
    \frac{\alpha^2}{4}\frac{1+\frac{c^2}{c_*^2}}{z_{\rm 
        max}-z_{\rm min}+4c^2\alpha^2}
  \]
  This linear fractional function of $c^2$ is increasing, decreasing,
  or constant if $\beta>0, \beta<0$ or $\beta=0$, respectively.
\end{proof}

In other words, $J_{\rm WRI}^{\alpha}$ has local minima far from the target
velocity $c_*$, in the same way as does $J_{\rm FWI}$. One of the
local minima will be the result of a local optimization almost surely,
unless the initial estimate of $c$ is ``within a wavelength'' of the
target velocity.

Note that $L$ is independent of $\alpha$ (definition \ref{eqn:defL},
and for small enough $\alpha$, $\beta > 0$ (definition
\ref{eqn:defbeta}. Conclude that the region $|c-c_*| > L\lambda$ is
independent of $L$, and the minimizer of $J_{\rm
  WRI}[c,\dl,\wl\delta(\cdot-z_s)]$ is $c=c_{\rm min}$ for small
enough $\alpha$. Therefore taking $\alpha$ small does not change the
multimodal nature of $J_{\rm WRI}^{\alpha}$: there remain multiple far-apart
local minima, no matter how small $\alpha$ may be.

\section{Discussion}
Theorems \ref{thm:thm1} and \ref{thm:thm2} call out the chief
conclusions of this work: that at least for the 1D acoustic
transmission inverse problem, both $J_{\rm FWI}[c,d,w]$ and
$J_{\rm WRI}^{\alpha}[c,d,q]$ exhibit local minima (in $c$) far from the global
minimum for consistent data, the domain of attraction of the
global minimizer can be arbitrarily small, and these properties
persist as $\alpha \rightarrow 0$. These are striking
conclusions, but the 1D acoustic transmission problem is very special
and lacks fidelity to field practice. I shall show how these
approaches to solving this special problem share properties with a
much larger family of inversion methods. The theory developed to
explain these properties suggests methods that may not suffer the
non-convexity of FWI and WRI, and in fact do not in several cases that
I will mention.

@@

One step in this construction involves building a constitutive defect
(pressure) source that is equivalent to a force (velocity) source, in the sense of generating the
same solution outside of the source support. This construction is
presented here in the context of the 1D acoustic problem, but is a
special case of a much more general construction of considerable
interest in its own right \cite[]{BurridgeKnopoff:64}.
